---
title: "Assignment 5"
author: "TQ Senkungu"
date: "6/14/2018"
output:
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Munging

###a. First I import the data and add column names.
```{r yob2016a}
setwd("~/Dropbox/Senkungu Fam/Education/SMU/Courses/MSDS 6306 Doing Data Science/GitHub Repo/smu-msds-homework/6306-homework/unit05")
df <- read.delim("yob2016.txt", header = FALSE, sep = ";")
names(df) <- c("Name", "Sex", "Frequency")
```
###b. This is the summary and structure
```{r yob2016b}
str(df)
summary(df)
```
###c. I list the misspelled name
```{r yob2016c}
misspelled_name <- df[grep("yyy",df$Name), ]
misspelled_name
```
###d. Here is the list with the missplled name removed
```{r yob2016d}
y2016 <- df[-c(216), ]
```

##2. Data Merging for the 2016 Data

###a. Import the file into R
```{r y2015a}
setwd("~/Dropbox/Senkungu Fam/Education/SMU/Courses/MSDS 6306 Doing Data Science/GitHub Repo/smu-msds-homework/6306-homework/unit05")
y2015 <- read.delim("yob2015.txt", header = FALSE, sep = ",")
names(y2015) <- c("Name", "Sex", "Frequency")
```

###b. Here are the last 10 rows.
```{r yob2015b}
tail(y2015, 10)
```
All of the last 10 items which are all names starting with Z were also used just 5 times. There are no NA values.

###c. I merged the two tables by name and sex and checked to ensure there are no NA values.
```{r y2015c}
final <- merge(y2015,y2016,by=c("Name","Sex"))
sapply(final, function(x) any(is.na(x)))
```


##3. Data Summary

###a. Total the frequencies of both years
```{r sum the frequencies}
final$Total <- rowSums(final[,c("Frequency.x", "Frequency.y")])
head(final)
```

I don't understand question 3a. But there were 26549 total people on the list. If we say these are popular names, that's the number.

###b. I sorted the data by the Total column.
```{r sort_totals}
final_sorted <- final[order(-final$Total), ]
head(final_sorted, 10)
```
Here are the names top 10 most popular names:

Emma
Olivia
Noah
Liam
Sophia
Ava
Mason
William
Jacob
Isabella

###c. Here is the list without boys
```{r remove boys}
final_sorted_female <- final_sorted[-which(final_sorted$Sex  == "M"),]
head(final_sorted_female,10)
```

###d. I created a csv with the names and frequencies.

```{r csv export}
final_girls_list <- head(final_sorted_female[,c("Name", "Total")], 10)
final_girls_list
write.csv(final_girls_list, file = "girl_names.csv", row.names=FALSE)
```

github location: https://github.com/tikisen/smu-msds-homework/tree/master/6306-homework/unit05