---
title: "MSDS 6306 Unit 04 Assignment"
author: "TQ Senkungu"
date: "6/6/2018"
output:
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get Five Thirty Eight Data

We get some data from the folks at Five Thirty Eight to do some analysis. The data is from a file on college_recent_grads. We pull it in and get some information about the file.

First I had to be sure to install the `fivethirtyeight` package and run it.

The article is linked at https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/


```{r college_recent_grads, cache=TRUE}
install.packages("fivethirtyeight", repos =  "http://cran.us.r-project.org")
library(fivethirtyeight)
df <- college_recent_grads
vignette("fivethirtyeight", package = "fivethirtyeight")
dim(df)
colnames(df)
```

## Look at Information on Major Categories

Here we get the number of columns of data then focus on the major categories. We create a plot of how many observations there are in each major category.

```{r major_count_plot}
length(colnames(df))
major_count <- table(df$major_category)
barplot(major_count, las=2, main = "Frequency of College Majors from Carnevale et al", xlab = "Major", ylab = "Frequency", horiz = TRUE, col = "red")
write.csv(df,row.names = FALSE, file = "/Users/tq/Dropbox/Senkungu Fam/Education/SMU/Courses/MSDS 6306 Doing Data Science/GitHub Repo/smu-msds-homework/6306-homework/unit04/college_recent_grads.csv")
```

##Github Repository

Here is a link to my homework Git repository

https://github.com/tikisen/smu-msds-homework 
