---
title: "Assignment 10"
author: "TQ Senkungu"
date: "7/18/2018"
output:
    html_document:
        keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mental Health Clinics

### b. State Abbreviations

```{r state abbreviations}
#load the experimental VA data
load("~/Dropbox/Senkungu Fam/Education/SMU/Courses/MSDS 6306 Doing Data Science/Homework/Assignment 10/N-MHSS-2015-DS0001-bndl-data-r/N-MHSS-2015-DS0001-data/N-MHSS-2015-DS0001-data-r.rda")

#Create a list of the unique state names
state_abbr <- unique(mh2015_puf$LST)
state_abbr
```

### c. Filter to only CONUS medical center count

```{r filter to conus count}
#load dplyr to do counts
library(dplyr)

#trim blank spaces in LST (state) column. It was jacking up the logical statement to subset the data frame
mh2015_puf$LST <- trimws(mh2015_puf$LST)

#create a subset of CONUS states
mh2015_puf_conus <- subset(mh2015_puf, !(LST %in% c("AK", "HI", "AS", "GU", "PR", "VI")))
#summary(mh2015_puf)

#make function to count occurences of something 
count_up <- function(x) {
  length(unique(x))
}

#use the function to count and aggregate the states
fac_per_state_conus <- aggregate(FUN = count_up, data=mh2015_puf_conus, . ~ LST)

#display a few of the states
head(fac_per_state_conus)

#isolate the state abbreviations and counts
fac_per_state_conus_t <- fac_per_state_conus[,c(1,2)]

#rename the facility count column
colnames(fac_per_state_conus_t)[2] <- "facility_count"

#verify the class of the final table
class(fac_per_state_conus_t)

#display the state count table
fac_per_state_conus_t
```

### d. Create ggplot of facilities

```{r ggplot of facilities}
#load ggplot library for barplot
library(ggplot2)

#load dplyr to count rows
library(dplyr)

#make states column a factor to create colors per state
fac_per_state_conus_t$LST <- as.factor(fac_per_state_conus_t$LST)
#verify class as factor
class(fac_per_state_conus_t$LST)

#make all main labels centered
theme_update(plot.title = element_text(hjust = 0.5))

#create plot dataset
va_plot <- ggplot(data = fac_per_state_conus_t, aes(x=LST, y=facility_count, fill = LST)) + geom_bar(stat = "identity")

#display plot, rotate state labels and add table and axis lables
va_plot + theme(axis.text.x=element_text(angle = 90, vjust = 0.5), legend.position="none") + scale_color_manual(values = LST) + labs(title="Number of VA Facilities by State", fill = "State", x= "State", y= "Facility Count")

#these were some color options I was playing with and will return to experiment with later
#library(RColorBrewer)
#library(wesanderson)
#va_plot <- ggplot(data = fac_per_state_conus_t, aes(x=LST, y=facility_count, fill = LST)) + geom_bar(stat = "identity")
#va_plot + theme(axis.text.x=element_text(angle = 90, vjust = 0.5), legend.position="top") +scale_color_manual(values=wes_palette(name="Royal1"))
#https://ggplot2.tidyverse.org/reference/scale_manual.html
#https://ggplot2.tidyverse.org/reference/scale_brewer.html
```
## Clean and Bring in New Features

### a. Read statesize.csv into file
```{r add new data}
statesize <- read.csv("~/Dropbox/Senkungu Fam/Education/SMU/Courses/MSDS 6306 Doing Data Science/Homework/Assignment 10/statesize.csv", header = TRUE)
fac_per_state_sizes <- merge(fac_per_state_conus_t, statesize, by.x = "LST", by.y = "Abbrev")
fac_per_state_sizes
```
I'm not sure what the question is referring to regarding the paste() function. All that I can see regarding any difference is that the state size table doesn't include DC. It may be a reference to the extra spaces in the LST column, but I stripped those out to make the subset in problem 1.

### b. Correct the state problem

I was able to merge the tables in part a. with no issues except the one stated regarding the stripped extra spaces.

### c. VA Hospitals per 1000 square miles

```{r hospitals per 1000 sq mi}
#check data type to see if we can do math
str(fac_per_state_sizes)

#calculation and field creation for facilities per 1000 sq miles
fac_per_state_sizes$state_fac_rate <- fac_per_state_sizes$facility_count/(fac_per_state_sizes$SqMiles/1000)

#display table
fac_per_state_sizes
```

### d. Plot of VAs per square thousand miles

```{r plot per thousand sq mi}
#create plot dataset
va_plot_sqft <- ggplot(data = fac_per_state_sizes, aes(x=reorder(LST, -state_fac_rate), y=state_fac_rate, fill = Region)) + geom_bar(stat = "identity")

#display plot, rotate state labels and add table and axis lables
va_plot_sqft + theme(axis.text.x=element_text(angle = 90, vjust = 0.5)) + scale_color_manual(values = Region) + labs(title="Number of VA Facilities per Thousand Sq Miles by State", fill = "State", x= "State", y= "Facility Count per 1000 Miles")
```

### e. Observations of Facility Rate by Region

The rate of facilities seems to be particularly high in the northeast. I'm sure that's mostly due to population density and relatively smaller state size. 

The south seems to be pretty much in the midrage with the exception of Maryland and Deleware which while maybe technically in the South would be considered more Northeast than South today.

The West seems to be on the low end of the spectrum for exactly the opposite reason. There are large states with spread out populations. For example, California has almost twice the population of New York, but more than 3 times the size. If the military populations occur at anything like the same rate, it would make sense that the facilities per 1000 sq ft would be higher in New York.

Before beginning modeling, I'd suggest that we also do a heat map of population density and overlay the VA locations. Even better would be an overlay of a military veterans location heatmap with VA locations. If we could combine this with a proxy for where the need is like wait times or utilization rates, that would drive where new facilities are more needed.

Then do a random forest because that seems to be the asnwer to every issue. :)